---
title: "Final Project"
author: "Group members: Allison Li, Youngjun Cho, and Nina Mo, Tutorial section:203, and Group number: 04"
subtitle: Analysis of Global Internet Usage
output:
  ioslides_presentation: default
  beamer_presentation: default
  slidy_presentation: default
  widescreen: yes
---


```{r, echo=FALSE, message=FALSE, warning=FALSE}
# echo=FALSE will stop the code chunk from appearing in the knit document
# warning=FALSE and message=FALSE will stop R messages from appearing in the knit document
library(tidyverse)
```
```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.height=3}
library(tidyverse)
path <- "https://raw.githubusercontent.com/ntaback/UofT_STA130/master/Fall2018/Finalproject/"

democracyindex2017 <- read_csv(paste0(path,"democracyindex2017.csv"))
education_cia2017 <- read_csv(paste0(path,"education_cia2017.csv"))
gdpppp_cia2017 <- read_csv(paste0(path,"gdpppp_cia2017.csv"))
lifeexpect_cia2017 <- read_csv(paste0(path,"lifeexpect_cia2017.csv"))
healthexpend_cia2017 <- read_csv(paste0(path,"healthexpend_cia2017.csv"))
internetusers_cia2017 <- read_csv(paste0(path,"internetusers_cia2017.csv"))
telephonelines_cia2017 <- read_csv(paste0(path,"telephonelines_cia2017.csv"))
population_cia2017 <- read_csv(paste0(path,"population_cia2017.csv"))
world_regions <- read_csv(paste0(path,"world_regions.csv"))

internet_usage <- inner_join(x= internetusers_cia2017, y = population_cia2017, by = "Country")%>%
  mutate(percentage_of_users = `INTERNET USERS`/POPULATION)%>%
  arrange(desc(percentage_of_users))%>%
  select(Country, `INTERNET USERS`, POPULATION, percentage_of_users)%>%
  filter(percentage_of_users <= 1)

library(tidyverse)
library(maps)
world <- map_data("world") 
#internetusers_cia2017 <- read_csv("internetusers_cia2017.csv")

iu <- internet_usage %>% rename(region = Country) 

iu$region[4] <- "USA" # to match world map data

iu <- semi_join(iu, world, by = "region")

map_internet <- ggplot()

map_internet <- map_internet + geom_map(
  data = world,
  map = world,
  aes(x = long, y = lat, map_id = region),
  fill = "#ffffff",
  color = "#ffffff",
  size = 0.20
  )
  
  map_internet <- map_internet + geom_map(
  data = iu,
  map = world,
  aes(fill = percentage_of_users, map_id = region),
  color = "#ffffff",
  size = 0.15
  )
  
  map_internet <- map_internet + scale_fill_continuous(low = 'thistle2', high = 'darkblue',
  guide = 'colorbar')
  
regions_internet_usage <- inner_join(x= internet_usage, y= world_regions, by= "Country")%>%
filter(Region != "Asic & Pacific")%>%
ggplot(aes(x= Region, y = percentage_of_users))+
geom_boxplot() + ggtitle("Internet Usage by Region") +
xlab("Region") + ylab("Internet Usage Rate")

gdp <- gdpppp_cia2017 %>% select(Country, 'GDP - PER CAPITA (PPP)')
gdp_internet <- inner_join(x = internet_usage, y = gdp , by = "Country")%>%
  mutate(GDP_PER_CAPITA = as.numeric(gsub('[$, ]','',`GDP - PER CAPITA (PPP)`)))
gdp_internet <- within(gdp_internet, rm(`GDP - PER CAPITA (PPP)`))

econ_graph <- ggplot(data=gdp_internet, aes(x= GDP_PER_CAPITA, y=percentage_of_users))+
  geom_point() + geom_smooth(method='lm') + ggtitle("The effect of Economy on Internet Usage") +
  xlab("GDP per Capita") + ylab("Internet Usage Rate")

life <- lifeexpect_cia2017 %>%
  select(Country, '(YEARS)')
life_internet <- inner_join(x= internet_usage, y= life, by = "Country")%>%
  mutate(years_group = rep(NA, 215))

life_internet$years_group[life_internet$'(YEARS)' <= 60] <- '< 60'
life_internet$years_group[life_internet$'(YEARS)' > 60 & life_internet$'(YEARS)' < 70] <- '60-70'
life_internet$years_group[life_internet$'(YEARS)' >= 70] <- '>70'

health_bp <- life_internet%>%ggplot(aes(x= years_group, y= percentage_of_users))+
  geom_boxplot() + ggtitle("The effect of Life Expectancy on Internet Usage") +
  xlab("Life Expectancy") + ylab("Internet Usage Rate")

health_sp <- ggplot(data = life_internet, aes(x = `(YEARS)` , y = percentage_of_users))+
  geom_point() + geom_smooth(method = 'lm') + ggtitle("The effect of Life Expectancy on Internet Usage") +
  xlab("Life Expectancy") + ylab("Internet Usage Rate")

dem <- democracyindex2017 %>%
  select(Country, Score, Category)
democracy <- inner_join(x = internet_usage, y = dem, by = "Country")%>%
  mutate(Score = as.numeric(Score))

# summary(lm(percentage_of_users ~ Category, data=democracy))$coefficients
dem_graph <- ggplot(data = democracy, aes(x = Score, y = percentage_of_users, colour = Category)) +
   geom_point(alpha=0.5) + geom_smooth(method = 'lm') + ggtitle("The effect of Democracy on Internet Usage") +
  xlab("Democracy Score") + ylab("Internet Userage")

education_internet <- inner_join(x = internet_usage, y = education_cia2017, by = 'Country')%>%
  select(Country, percentage_of_users, `(% OF GDP)`)

education_gdp <- inner_join(x = education_internet, y = gdpppp_cia2017, by = 'Country')%>%
  mutate(GDP_PER_CAPITA = as.numeric(gsub('[$, ]','',`GDP - PER CAPITA (PPP)`)))%>%
  mutate(edu_gdp = (`(% OF GDP)` * GDP_PER_CAPITA))
education_gdp <- within(education_gdp, rm(Rank, `Date of Information`, `GDP - PER CAPITA (PPP)`))

edu_graph <- ggplot(data = education_gdp, aes(x = edu_gdp, y = percentage_of_users))+
         geom_point() + geom_smooth(method='lm') + ggtitle("The effect of Education on Internet Usage") +
  xlab("Education Budget") + ylab("Internet Userage")
```

## Introduction - What is the internet
- The internet is a system of global networks that transmits data and connect devices globally
  - It is a network of networks
  
  - Invented in the 1960s, this technological phenomennon quickly rose to popularity and now dominates     households, schools, and workplaces. 
  
  - Its prevalence in our everyday life and our reliance on it prompts us to question what affects our     society's usage of internet. 

## Introduction - Cont.
For our purposes, we will be using the data sets from the CIA world fact book to compare each country

  - Internet user: "compares the number of users within a country that access the Internet"
  
  - Population: "compares estimates from the US Bureau of the Census based on statistics from
    population censuses, vital statistics registration systems, or sample surveys pertaining to the       recent past and on assumptions about future trends"

## Objectives
In this presentation we will:

- Q1: Define how we measure internet usage in a country

- Q2: Compare internet usage within regions 

- Q3: Define concepts such as democracy, education, economy, and health in order to assess their impact on internet usage

## Q1 - How do we define internet usage in a country?
  - Each country has a different population. Thus, to ensure that our comparisons are not affected by population, we defined internet usage rate as: 
  
    - number of internet users / population

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.height=3}
glimpse(internet_usage)
```

## Q2 - Internet usage by regions of the world: World Map

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.height=3}
map_internet
```



## Q2 - Internet use by regions of the world: Boxplot 

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.height=3}
regions_internet_usage 
```

## Q3 - How we defined Economy:
  - Defined the economy of a country through their GDP per Capita
  - What is GDP per Capita?
```{r}
glimpse(gdp_internet)
```
  

## Q3 - Impact of Economy on Internet Use:

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.height=3}
econ_graph
```

## Q3 - How we defined Health:
  - Defined the health of a country through the life expectancy of their citizens

```{r}
glimpse(life_internet)
```

## Q3 - Impact of Health on Internet Use:

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.height=3}
health_sp
```

## Q3 - How we defined Democracy:
  - We defined the democracy of a country based on the score that they received
  - What does the score represent?
```{r}
glimpse(democracy)
```


## Q3 - Impact of Democracy on Internet Usage:
  - Coloured based on type of government

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.height=3}
dem_graph
```



## Q3 - How we defined Education

  - We defined the education based on how much money the governent spent on education
  
  - To adjust for how countries have different sizes of GDP:
  
    - We multiplied the % of GDP spent on education with the country's GDP

```{r}
glimpse(education_gdp)
```

## Q3 - Impact of Education on Internet Usage:

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.height=3}
edu_graph
```

## Q3 - Conclusion

     - Summary
  There is a positive correlation between the percentage of internet users and country's democracy score, education budget, life expectancy, and GDP. 
  
  Regions with the relatively higher median percentage of internet users include more countries with higher GDP and other three factors 
      
      - Concern
  Date of information is different in different data
  
  Is GDP a confounding variable?
  
  
